R Notebook

### 2022-03-22 src/main/r/biomarker.Rmd
update analysis

## explore other statistical models
model2 <- glmer(formula = 'case_or_control ~ gender + race + age + is_observed + (1 + is_observed | is_tested)', data = df_model, family = binomial(link = 'logit') )
summary(model2)
```

@@

plan(multisession, workers = 6)

screen_results2 <- phenotype_features %>% 
  #filter(termid == 'HP:0001880') %>%
  #filter(lab_date_rel_bracket == '(-1,0]') %>%
  group_by(lab_date_rel_bracket, termid) %>%
  nest() %>%
  #calling ungroup() is critical for furrr to work in parallel
  ungroup() %>%
  mutate(coefficient = future_map(data, fit_model, df_confounders_outcome = confounders_outcome)) %>%
  select(-data) %>%
  unnest(coefficient)

>>>>

## explore other statistical models
# model2 <- glmer(formula = 'case_or_control ~ gender + race + age + is_observed + (1 + is_observed | is_tested)', data = df_model, family = binomial(link = 'logit') )
# summary(model2)
```

@@

plan(multisession, workers = 8)

screen_results <- phenotype_features %>% 
  #filter(termid == 'HP:0001880') %>%
  #filter(lab_date_rel_bracket == '(-1,0]') %>%
  group_by(lab_date_rel_bracket, termid) %>%
  nest() %>%
  #calling ungroup() is critical for furrr to work in parallel
  ungroup() %>%
  mutate(coefficient = future_map(data, fit_model, df_confounders_outcome = confounders_outcome)) %>%
  select(-data) %>%
  unnest(coefficient)

### 2022-03-22 src/main/r/numeric_vs_hpo_for_CRC.Rmd
update analysis

# cache directory to read from or write to
CACHE_DIR <- './cache'
if (!dir.exists(CACHE_DIR)){
  dir.create(CACHE_DIR)
}
```

@@

Retrieve diagnoses information for the cohort.
```{r}
sql <- "SELECT * FROM cohort"
df_cohort <- get_df_from_db_or_cache(cache_dir = CACHE_DIR, file_name = 'colorectal_cancer_diagnoses.csv', db_access_callback = dbGetQuery, conn = dsca, statement = sql, force_rerun = FALSE)

@@

Retrieve related lab tests for the cohort. 
```{r}
sql <- read_file('../sql/get_labs_for_colorectal_cancer.sql')
df_labs <- get_df_from_db_or_cache(cache_dir = CACHE_DIR, file_name = "labs_colorectal_cancer.csv", db_access_callback = dbGetQuery, conn=dsca, statement = sql, force_rerun = FALSE)
df_labs %>% nrow()
```

@@

p1 <- plot_lab_raw_values_vs_rel_date(lab_loinc = '6690-2', ylim = c(0, 50)) + ggtitle('Leukocytes')

@@

```{r max.output=-1, message=FALSE}
require(furrr)
plan(multisession, workers = 4)

crc_analysis <- df_labs_and_outcomes %>%
  group_by(loinc) %>%
  nest() %>%
  mutate(result = future_map(data, pipeline_compare_for_outcome_prediction)) %>%
  select(-data) %>%
  unnest(result)

@@

```{r}
sql <- read_file('../sql/colorectal_cancer_cohort_demographics.sql')
df_case_demographics <- get_df_from_db_or_cache(cache_dir = CACHE_DIR, file_name = 'demographics_colorectal_cancer_cases.csv', db_access_callback = dbGetQuery, conn = dsca, statement = sql, force_rerun = FALSE)

@@

sql <- 'SELECT * FROM control_cohort'
df_control_demographics <- get_df_from_db_or_cache(cache_dir = CACHE_DIR, file_name = 'demographics_colorectal_cancer_controls.csv', db_access_callback = dbGetQuery, conn = dsca, statement = sql, force_rerun = FALSE)

sql <- read_file('../sql/get_labs_for_colorectal_cancer_control.sql')
df_control_labs <- get_df_from_db_or_cache(cache_dir = CACHE_DIR, file_name = 'labs_colorectal_cancer_sampled_controls.csv', db_access_callback = dbGetQuery, conn = dsca, statement = sql, force_rerun = FALSE)
```

@@

# labs for controls
df_labs_and_outcomes_controls <- df_control_labs %>% filter(!(loinc %in% c('21000-5', '32207-3'))) %>%
  as_tibble() %>%
  #filter(medical_record_number == '7007952')  %>%
  #mutate(lab_date = as.Date(lab_date)) %>%
  arrange(medical_record_number, loinc, desc(lab_date)) %>%
  group_by(medical_record_number, loinc) %>%
  mutate(sequence = row_number()) %>%
  ungroup() %>%
  filter(sequence == 1) %>%
  mutate(outcome = 'normal')

@@

crc_case_matched_controls_analysis
```

@@

crc_case_matched_controls_analysis_with_demographics <- df_labs_and_outcomes_case_matched_control %>%
  left_join(df_demographics_case_matched_control, by = 'medical_record_number') %>%
  mutate(age = as.duration(interval(date_of_birth, lab_date)) / dyears()) %>%
  filter(!(race %in% c('African American', 'American Indian or Alaska Native', 'Native Hawaiian or Other Pacific Islander'))) %>%
  # mutate(gender = as.factor(gender),
  #        race = as.factor(race)) %>% 
  # mutate(gender = relevel(gender, 'Male'),
  #        race = relevel(race, 'White')) %>%
  group_by(loinc) %>%
  nest() %>%
  mutate(result = future_map(data, pipeline_compare_for_outcome_prediction, include_demographics=TRUE, .options = furrr_options(seed = T))) %>%
  select(-data) %>%
  unnest(result)

>>>>

# cache directory to read from or write to
CACHE_DIR <- './cache'
if (!dir.exists(CACHE_DIR)){
  dir.create(CACHE_DIR)
}
FORCE_RERUN <- TRUE
```

@@

Retrieve diagnoses information for the cohort.
```{r}
sql <- "SELECT * FROM cohort"
df_cohort <- get_df_from_db_or_cache(cache_dir = CACHE_DIR, file_name = 'colorectal_cancer_diagnoses.csv', db_access_callback = dbGetQuery, conn = dsca, statement = sql, force_rerun = FORCE_RERUN)

@@

Retrieve related lab tests for the cohort. 
```{r}
sql <- read_file('../sql/get_labs_for_colorectal_cancer.sql')
df_labs <- get_df_from_db_or_cache(cache_dir = CACHE_DIR, file_name = "labs_colorectal_cancer.csv", db_access_callback = dbGetQuery, conn=dsca, statement = sql, force_rerun = FORCE_RERUN)
df_labs %>% nrow()
```

@@

p1 <- plot_lab_raw_values_vs_rel_date(lab_loinc = '26464-8', ylim = c(0, 50)) + ggtitle('Leukocytes')

@@

```{r max.output=-1, message=FALSE}
require(furrr)
plan(multisession, workers = 5)

crc_analysis <- df_labs_and_outcomes %>%
  group_by(loinc) %>%
  nest() %>%
  ungroup() %>%
  mutate(result = future_map(data, pipeline_compare_for_outcome_prediction)) %>%
  select(-data) %>%
  unnest(result)

@@

```{r}
sql <- read_file('../sql/colorectal_cancer_cohort_demographics.sql')
df_case_demographics <- get_df_from_db_or_cache(cache_dir = CACHE_DIR, file_name = 'demographics_colorectal_cancer_cases.csv', db_access_callback = dbGetQuery, conn = dsca, statement = sql, force_rerun = FORCE_RERUN)

@@

sql <- 'SELECT * FROM control_cohort'
df_control_demographics <- get_df_from_db_or_cache(cache_dir = CACHE_DIR, file_name = 'demographics_colorectal_cancer_controls.csv', db_access_callback = dbGetQuery, conn = dsca, statement = sql, force_rerun = FORCE_RERUN)

sql <- read_file('../sql/get_labs_for_colorectal_cancer_control.sql')
df_control_labs <- get_df_from_db_or_cache(cache_dir = CACHE_DIR, file_name = 'labs_colorectal_cancer_sampled_controls.csv', db_access_callback = dbGetQuery, conn = dsca, statement = sql, force_rerun = FORCE_RERUN)
```

@@

# labs for controls
df_labs_and_outcomes_controls <- df_control_labs %>% filter(!(loinc %in% c('21000-5', '32207-3'))) %>%
  as_tibble() %>%
  mutate(medical_record_number = as.character(medical_record_number)) %>% 
  mutate(local_test_code = as.character(local_test_code)) %>% 
  #filter(medical_record_number == '7007952')  %>%
  #mutate(lab_date = as.Date(lab_date)) %>%
  arrange(medical_record_number, loinc, desc(lab_date)) %>%
  group_by(medical_record_number, loinc) %>%
  mutate(sequence = row_number()) %>%
  ungroup() %>%
  filter(sequence == 1) %>%
  mutate(outcome = 'normal')

@@

crc_case_matched_controls_analysis %>% write.csv(file = 'cache/crc_case_matched_controls_analysis.csv', row.names = FALSE)

crc_case_matched_controls_analysis

```

@@

crc_case_matched_controls_analysis_with_demographics <- df_labs_and_outcomes_case_matched_control %>%
  left_join(df_demographics_case_matched_control, by = 'medical_record_number') %>%
  mutate(age = as.duration(interval(date_of_birth, lab_date)) / dyears()) %>%
  filter(!(race %in% c('African American', 'American Indian or Alaska Native', 'Native Hawaiian or Other Pacific Islander'))) %>%
  # mutate(gender = as.factor(gender),
  #        race = as.factor(race)) %>% 
  # mutate(gender = relevel(gender, 'Male'),
  #        race = relevel(race, 'White')) %>%
  group_by(loinc) %>%
  nest() %>%
  ungroup() %>% 
  mutate(result = future_map(data, pipeline_compare_for_outcome_prediction, include_demographics=TRUE, .options = furrr_options(seed = T))) %>%
  select(-data) %>%
  unnest(result)
